Sample-Based Determination of Angular Observables

### 2014-06-23 draft.tex
[draft] Corrections on first page

\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{xcolor}

@@

%% Draft Macros %%
\newcommand{\todo}[1]{{\color{red}\bf ToDo: #1}}
\newcommand{\danny}[1]{{\color{purple}#1}}
\newcommand{\fred}[1]{{\color{gold!85!black}#1}}
\newcommand{\nico}[1]{{\color{green!85!black}#1}}
\newcommand{\citeneeded}{{\color{red}\bf [cite needed]}}

@@

\preprint{SI-HEP-2014-xx}
\title{Sample Based Determination of Angular Observables}
\author{Frederik Beaujean}
\email{frederik.beaujean@lmu.de}
\affiliation{C2PAP, Universe Cluster, Ludwig-Maximilians-Universit\"at M\"unchen, Garching, Germany}
\author{Marcin Chrzaszcz}
\email{chrzaszcz.marcin@gmail.com}
\author{Nicola Serra}
\email{nicola.serra@cern.ch}
\affiliation{Physik-Institut, Universit\"at Z\"urich, Z\"urich, Switzerland}
\author{Danny van Dyk}
\email{vandyk@tp1.physik.uni-siegen.de}
\affiliation{Theoretische Physik 1, Naturwissenschaftlich-Technische Fakult\"at,
Universit\"at Siegen, Siegen, Germany}

\begin{abstract}
We study means to determine angular observables for a generic angular distribution
of either decays or scattering processes. \dots
\end{abstract}

@@

In this letter we study how to obtain a set of angular observables $P_i$
that arise in a generic multi-body processes without the need to carry out
an explicit fit of an angular distribution to data. Instead, we present a method
that only relies on orthogonality of angular functions, and estimation of integrals
by means of Monte Carlo techniques.\\

The initial motivation for studying what we wish to call the \emph{Method of Moments}
is the determination of angular observables in the decay rare, FCNC-mediated decay
$\bar{B}\to \bar{K}^*(\to \bar{K}\pi)\ell^+\ell^-$. However, we emphasize that
he method we describe in the remainder of this letter is applicable to arbitrary
decay or scattering processes.\\

For the purpose of this work, let $P_i$ be a set of functions of $m$ non-angular
kinematic variables $\nuvec$, $P_i \equiv P_i(\nuvec)$, with $\nuvec = (\nu_1, \dots \nu_m)$.
These include, among others, invariant masses or center of mass energies pertinent
to the respective decay or scattering processes. Let us further assume the angular
observables to be defined via a suitably factorizing probability distribution (PDF),
\begin{align}
    \label{eq:def-P}
    P(\nuvec, \thvec) = \sum_k P_i(\nuvec) \times f_i(\thvec)\,,
\end{align}
Here, the dependence on the $n$ decay angles $\thvec = (\theta_1, \dots, \theta_n)$ has been
explicitly factored out in terms of the angular functions $f_i(\thvec)$, which we assume to fulfill the
orthonormality relations
\begin{equation}
    \label{eq:def-ortho-rel}
    \int_\Omega f_i(\thvec) f_j(\thvec) \dd^n \theta = \delta_{ij}\,.
\end{equation}
(For the purpose of this letter, however, it suffices that the system of angular
functions $f_i(\thvec)$ can be transformed into an orthonorml basis. The transformations to these
bases for a selection of decays has been written out in the appendices \ref{app:btokll} through
\ref{app:lambdabtolambdall}.)\\

For particle decays, we $P$ is expressed in terms of the fully differential decay width,
\begin{align}
    \label{eq:def-P-decay}
    P(\nuvec, \thvec) \equiv \frac{1}{\Gamma}\frac{\dd^{n+m}\Gamma}{\dd \nu_1 \dots \dd \nu_m\, \dd \theta_1 \dots \theta_n}\,,
\end{align}
where $\Gamma$ is the total decay width. For a scattering process, one can similarly use
\begin{align}
    \label{eq:def-P-scattering}
    P(\nuvec, \thvec) \equiv \frac{1}{\sigma}\frac{\dd^{n+m}\sigma}{\dd \nu_1 \dots \dd \nu_m\, \dd \theta_1 \dots \theta_n}\,,
\end{align}
where the total cross section $\sigma$ is used for the normalisation. Since the determination of
the total decay width or total cross section can be quite difficult, we emphasize that different normalizations for
$P$ can be used. For instance, the total decay width (or cross section) of the process of interest can be replaced
by the corresponding quantity of a control-channel process.\\

@@

\section{Sample Based Determination}
\label{sec:sample-based-det}

>>>>

\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[svgnames]{xcolor}

@@

%% Draft Macros %%
\newcommand{\todo}[1]{{\color{red}\bf ToDo: #1}}
\newcommand{\danny}[1]{{\color{purple}#1}}
\newcommand{\fred}[1]{{\color{brown!85!black}#1}}
\newcommand{\nico}[1]{{\color{green!85!black}#1}}
\newcommand{\citeneeded}{{\color{red}\bf [cite needed]}}

@@

\preprint{SI-HEP-2014-xx}
\title{Sample-Based Determination of Angular Observables}
\author{Frederik Beaujean}
\email{frederik.beaujean@lmu.de}
\affiliation{C2PAP, Universe Cluster, Ludwig-Maximilians-Universit\"at M\"unchen, Garching, Germany}
\author{Marcin Chrzaszcz}
\email{chrzaszcz.marcin@gmail.com}
\author{Nicola Serra}
\email{nicola.serra@cern.ch}
\affiliation{Physik-Institut, Universit\"at Z\"urich, Z\"urich, Switzerland}
\author{Danny van Dyk}
\email{vandyk@tp1.physik.uni-siegen.de}
\affiliation{Theoretische Physik 1, Naturwissenschaftlich-Technische Fakult\"at,
Universit\"at Siegen, Siegen, Germany}

\begin{abstract}
We study means to determine angular observables for a generic angular distribution
of a decay or scattering process. \dots
\end{abstract}

@@

In this letter we study how to obtain a set of angular observables
$\{P_i\}$ that arise in a generic multi-body process without the need
to carry out a likelihood fit of an angular distribution to the
measured events. Instead, we present a method that only relies on
orthogonality of angular functions, and estimation of integrals
by means of Monte Carlo techniques. \fred{Paragraph sounds like abstract}\\

The initial motivation for studying what we wish to call the \emph{Method of Moments}
is the determination of angular observables in the rare decay FCNC-mediated decay
$\bar{B}\to \bar{K}^*(\to \bar{K}\pi)\ell^+\ell^-$. However, we emphasize that
he method we describe in the remainder of this letter is general and applies to arbitrary
decay or scattering processes.\\

Let $\thvec = (\theta_1, \dots \theta_n)$ denote the set of all
angles, and let $\nuvec = (\nu_1, \dots \nu_m)$ denote the set of all
other non-angular kinematic variables needed to fully specify the
final state of the process under study. For example, $\nuvec$ may
include invariant masses or center-of-mass energies. We define an
angular observable $P_i$ as a coefficient in the probability density
function (PDF) of the process,
\begin{align}
    \label{eq:def-P}
    P(\nuvec, \thvec) = \sum_i P_i(\nuvec) \times f_i(\thvec)\,,
\end{align}
Here, the dependence on the $n$ decay angles $\thvec$ has been
explicitly factored out in terms of the angular functions
$f_i(\thvec)$, which we assume to fulfill the orthonormality relations
\begin{equation}
    \label{eq:def-ortho-rel}
    \int_\Omega f_i(\thvec) f_j(\thvec) \dd^n \theta = \delta_{ij}\,.
\end{equation}
(For the purpose of this letter, however, it suffices that the system
of angular functions $f_i(\thvec)$ can be transformed into an
orthonormal basis. The transformation needs to be worked out on a
case-by-case basis. For a selection of $B$ decays, the transformations
are listed in the appendices \ref{app:btokll} through
\ref{app:lambdabtolambdall}.)\\

For particle decays, $P$ is generally expressed in terms of the fully differential decay width,
\begin{align}
    \label{eq:def-P-decay}
    P(\nuvec, \thvec) \equiv \frac{1}{\Gamma}\frac{\dd^{n+m}\Gamma}{\dd \nu_1 \dots \dd \nu_m\, \dd \theta_1 \dots \theta_n}\,,
\end{align}
where $\Gamma$ is the total decay width. For a scattering process, one can similarly use
\begin{align}
    \label{eq:def-P-scattering}
    P(\nuvec, \thvec) \equiv \frac{1}{\sigma}\frac{\dd^{n+m}\sigma}{\dd \nu_1 \dots \dd \nu_m\, \dd \theta_1 \dots \theta_n}\,,
\end{align}
where the total cross section $\sigma$ is used for the normalisation. Since the determination of
the total decay width or total cross section can be quite difficult, we emphasize that different normalizations for
$P$ can be used. For instance, the total decay width (or cross section) of the process of interest can be replaced
by the corresponding quantity of a control-channel process.\\

@@

\section{Sample-Based Determination}
\label{sec:sample-based-det}
